---
title: "p8105_hw6_fl2714"
author: "Fangchi"
date: "2024-11-26"
output: github_document
---

```{r}

library(tidyverse)
library(modelr)
library(mgcv)
library(dplyr)
library(ggplot2)
library(purrr)
library(tidyr)
library(rsample)
library(car)

```

### Problem 2

```{r}

# load data
url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
homicide_data <- read.csv(url, stringsAsFactors = FALSE)

```


```{r}

cleaned_data <- homicide_data %>%
  mutate(city_state = paste(city, state, sep = ", ")) %>% # create variable
  mutate(solved_binary = ifelse(disposition == "Closed by arrest", 1, 0)) %>% # whether the homicide is solved
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>%
  filter(victim_race %in% c("White", "Black")) %>% # limit victim_race
  mutate(
    victim_age = ifelse(victim_age == "Unknown", NA, victim_age),  
    victim_age = as.numeric(victim_age)                            
  ) %>% # victim_age is numeric
  drop_na(victim_age)  

```



```{r}

baltimore_data <- cleaned_data %>%
  filter(city_state == "Baltimore, MD")

glm_baltimore <- glm(solved_binary ~ victim_age + victim_sex + victim_race,
                     data = baltimore_data, family = "binomial")

glm_baltimore |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(OR = exp(estimate), 
         CI_low = exp(conf.low), 
         CI_high = exp(conf.high)) |>  
  select(term, log_OR = estimate, OR, p.value, CI_low, CI_high) |> 
  knitr::kable(digits = 3)

```

The estimate for the adjusted odds ratio comparing male victims to female victims is 0.426, with a 95% confidence interval of [0.324, 0.558], indicating that the odds of solving homicides for male victims are significantly lower than those for female victims, controlling for other variables.

```{r}

# Group the data by city_state
results <- cleaned_data %>%
  group_by(city_state) %>% 
  nest() %>%  # Create a nested data frame for each city
  mutate(
    glm_model = map(data, ~ glm(solved_binary ~ victim_age + victim_sex + victim_race,
                                data = .x, family = "binomial")),  # Fit the logistic model for each city
    glm_tidy = map(glm_model, ~ broom::tidy(.x, conf.int = TRUE))  # Extract coefficients and CI
  ) %>%
  unnest(cols = glm_tidy) %>%  # Unnest the coefficients into a flat data frame
  filter(term == "victim_sexMale") %>%  # Select only the row for male vs female comparison
  mutate(
    OR = exp(estimate),  # Convert log(OR) to OR
    CI_low = exp(conf.low),  # Convert lower CI bound
    CI_high = exp(conf.high)  # Convert upper CI bound
  ) %>%
  select(city_state, OR, CI_low, CI_high, p.value)  # Keep relevant columns

# Display results
results %>%
  knitr::kable(digits = 3)


```
```{r}


# Create the data for plotting
plot_data <- results %>%  
  arrange(OR) %>%  # Arrange cities by OR
  mutate(city_state = factor(city_state, levels = city_state))  

# Plot the ORs and CIs using ggplot
ggplot(plot_data, aes(x = city_state, y = OR)) +
  geom_point(size = 3) +  
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = 0.2) + 
  coord_flip() +  # Flip coordinates for better readability
  theme_minimal() +  # Use a minimal theme
  labs(
    title = "Adjusted Odds Ratios (OR) and Confidence Intervals (CI) by City",
    x = "City",
    y = "Adjusted Odds Ratio (OR)"
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at OR = 1
  theme(
    axis.text.y = element_text(size = 8), 
    axis.text.x = element_text(size = 8)
  )

```
General Trend: In most cities, male victims are less likely to have their homicides solved compared to female victims (OR < 1).
Statistical Significance: Cities like Baltimore, MD and Atlanta, GA show significant effects with confidence intervals entirely below 1, indicating a strong association.
Variability: Some cities show no significant difference (OR ≈ 1), while a few have OR > 1 with wide confidence intervals, suggesting uncertainty or smaller sample sizes.

### Problem 3

```{r}
# Load and clean the data

birthweight <- read.csv("data/birthweight.csv")

birthweight <- birthweight %>%
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
    frace = factor(frace),
    mrace = factor(mrace)
  ) %>%
  drop_na()  

```


```{r}

model_hypothesis <- lm(bwt ~ babysex + bhead + blength + gaweeks + wtgain + smoken, data = birthweight)
summary(model_hypothesis)

```

This model is based on hypothesized factors known to influence birthweight, supported by biological and clinical evidence:

babysex: Male babies typically have higher birthweight than female babies
bhead (head circumference): A direct indicator of the baby’s size
blength (length): size-related factor correlated with birthweight
gaweeks (gestational age): Longer gestation often leads to higher birthweight
wtgain (maternal weight gain): Reelated to maternal health and nutrition during pregnancy
smoken (maternal smoking): Smoking restricts fetal growth

These variables were chosen based on domain knowledge about the factors influencing birthweight.

```{r}

vif(model_hypothesis)

```


```{r}

birthweight <- birthweight %>%
  add_predictions(model_hypothesis, var = "pred") %>%
  add_residuals(model_hypothesis, var = "resid")

ggplot(birthweight, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

```
```{r}

model_main_effects <- lm(bwt ~ blength + gaweeks, data = birthweight)
summary(model_main_effects)

```
```{r}

model_interactions <- lm(bwt ~ bhead * blength * babysex, data = birthweight)
summary(model_interactions)

```
```{r}

set.seed(123)  
cv_df <- crossv_mc(birthweight, 100) %>%  
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df <- cv_df %>%
  mutate(
    model_hypothesis = map(train, ~ lm(bwt ~ babysex + bhead + blength + gaweeks + wtgain + smoken, data = .)),
    model_main_effects = map(train, ~ lm(bwt ~ blength + gaweeks, data = .)),
    model_interactions = map(train, ~ lm(bwt ~ bhead * blength * babysex, data = .)),
    rmse_hypothesis = map2_dbl(model_hypothesis, test, ~ rmse(model = .x, data = .y)),
    rmse_main_effects = map2_dbl(model_main_effects, test, ~ rmse(model = .x, data = .y)),
    rmse_interactions = map2_dbl(model_interactions, test, ~ rmse(model = .x, data = .y))
  )

```


```{r}

cv_long <- cv_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  )


ggplot(cv_long, aes(x = model, y = rmse)) +
  geom_violin() +
  labs(title = "Cross-Validated RMSE for Each Model",
       x = "Model",
       y = "RMSE") +
  theme_minimal()

```

